Wallace Session 2023-06-26
================

Please find below the R code history from your *Wallace* v2.0.5 session.

You can reproduce your session results by running this R Markdown file
in RStudio.

Each code block is called a “chunk”, and you can run them either
one-by-one or all at once by choosing an option in the “Run” menu at the
top-right corner of the “Source” pane in RStudio.

For more detailed information see <http://rmarkdown.rstudio.com>).

### Package installation

Wallace uses the following R packages that must be installed and loaded
before starting.

```{r}
library(spocc)
library(spThin)
library(dismo)
library(rgeos)
library(ENMeval)
library(wallace)
```

The *Wallace* session code .Rmd file is composed of a chain of module
functions that are internal to *Wallace*. Each of these functions
corresponds to a single module that the user ran during the session. To
see the internal code for these module functions, click on the links in
the .Rmd file. Users are encouraged to write custom code in the .Rmd
directly to modify their analysis, and even modify the module function
code to further customize. To see the source code for any module
function, just type its name into the R console and press Return.

```{r}
# example:
# just type the function name and press Return to see its source code
# paste this code into a new script to edit it
occs_queryDb
```

Your analyses are below.

------------------------------------------------------------------------

## Analysis for *Carcharhinus perezi* (Cp)

### Obtain Occurrence Data

You searched the gbif database for *Carcharhinus perezi*, limited to
5000 records. You decided to remove occurrences without uncertainty
information? FALSE

```{r}
# Query selected database for occurrence records
queryDb_Cp <- occs_queryDb(
  spNames = "Carcharhinus perezi", 
  occDb = "gbif", 
  occNum = 5000,
  RmUncertain = FALSE)
occs_Cp <- queryDb_Cp$Carcharhinus_perezi$cleaned
```

### Obtain environmental data

Using user-specified variables.

```{r}
## Specify the directory with the environmental variables
dir_envs_Cp <- "D:/Wallace/GIT/MARSPECTca"
envs_path <- file.path(dir_envs_Cp, c('b1_EastWestAspect.tif', 'b2_NorthSouthAspect.tif', 'b3_PlanCurvature.tif', 'b4_ProfileCurvature.tif', 'b5_DistancetoShore.tif', 'b6_BathymetricSlope.tif', 'b7_Concavity.tif', 'b8_MeanAnnualSSS.tif', 'b9_MinimumMonthlySSS.tif', 'b10_MaximumMonthlySSS.tif', 'b11_AnnualRangeSSS.tif', 'b12_AnnualVarianceSSS.tif', 'b13_MeanAnnualSST.tif', 'b14_SSTcoldestIceFreeMonth.tif', 'b15_SSTwarmestIceFreeMonth.tif', 'b16_AnnualRangeSST.tif', 'b17_AnnualVarianceSST.tif', 'bathymetry_30s.tif'))
# Create environmental object 
envs_Cp <- envs_userEnvs(
  rasPath = envs_path,
  rasName = c('b1_EastWestAspect.tif', 'b2_NorthSouthAspect.tif', 'b3_PlanCurvature.tif', 'b4_ProfileCurvature.tif', 'b5_DistancetoShore.tif', 'b6_BathymetricSlope.tif', 'b7_Concavity.tif', 'b8_MeanAnnualSSS.tif', 'b9_MinimumMonthlySSS.tif', 'b10_MaximumMonthlySSS.tif', 'b11_AnnualRangeSSS.tif', 'b12_AnnualVarianceSSS.tif', 'b13_MeanAnnualSST.tif', 'b14_SSTcoldestIceFreeMonth.tif', 'b15_SSTwarmestIceFreeMonth.tif', 'b16_AnnualRangeSST.tif', 'b17_AnnualVarianceSST.tif', 'bathymetry_30s.tif'),
  doBrick = FALSE)
occs_xy_Cp <- occs_Cp[c('longitude', 'latitude')]
occs_vals_Cp <- as.data.frame(raster::extract(envs_Cp, occs_xy_Cp, cellnumbers = TRUE))
# Remove duplicated same cell values
occs_Cp <- occs_Cp[!duplicated(occs_vals_Cp[, 1]), ]
occs_vals_Cp <- occs_vals_Cp[!duplicated(occs_vals_Cp[, 1]), -1]
# remove occurrence records with NA environmental values
occs_Cp <- occs_Cp[!(rowSums(is.na(occs_vals_Cp)) >= 1), ]
# also remove variable value rows with NA environmental values
occs_vals_Cp <- na.omit(occs_vals_Cp)
# add columns for env variable values for each occurrence record
occs_Cp <- cbind(occs_Cp, occs_vals_Cp)
```

### Process Occurrence Data

Remove occurrences outside of user drawn polygon

```{r}
occs_Cp <- poccs_selectOccs(
  occs = occs_Cp,
  polySelXY = matrix(c(-79.168605, -82.685205, -82.575312, -79.322456, -75.564089, -71.080423, -69.036399, -66.596757, -63.124114, -63.431817, -64.069201, -64.420861, -65.387926, -70.706784, -74.443172, -77.652071, -78.487263, -79.168605, 16.782519, 14.39112, 11.414408, 10.400365, 11.048027, 12.767942, 13.003554, 11.479015, 11.608185, 15.600883, 17.181795, 17.977753, 17.873223, 17.47545, 17.517362, 17.496407, 17.47545, 16.782519), ncol = 2, byrow = FALSE),
  polySelID = 1607)
```

### Process environmental data

Sampling of 100 background points and corresponding environmental data
using a user drawn background extent with a 0 degree buffer.

```{r}
# Create a background extent based on user drawn polygon
bgExt_Cp <- penvs_drawBgExtent(
  polyExtXY = matrix(c(-80.071559, -82.665052, -82.665052, -81.170497, -78.401174, -78.168427, -77.662916, -76.135392, -75.739775, -75.080412, -74.497975, -73.662782, -72.585823, -71.607769, -69.434351, -68.060679, -65.115526, -62.56599, -61.51101, -61.693613, -62.572763, -64.111276, -68.023494, -71.408222, -74.79295, -78.177679, -79.452446, -80.071559, 16.928734, 14.815379, 11.262592, 9.75034, 9.793648, 9.426879, 9.144978, 9.719378, 10.573716, 11.134782, 11.231781, 11.382604, 12.038817, 12.671994, 12.822012, 11.705526, 10.951473, 11.350292, 13.656162, 16.043834, 17.51638, 17.976774, 17.264761, 17.432545, 17.264761, 17.264761, 17.432545, 16.928734),ncol=2,byrow=FALSE), 
  polyExtID = 6923, 
  drawBgBuf = 0, 
  occs = occs_Cp)
# Mask environmental data to provided extent
bgMask_Cp <- penvs_bgMask(
  occs = occs_Cp,
  envs = envs_Cp,
  bgExt = bgExt_Cp)
# Sample background points from the provided area
bgSample_Cp <- penvs_bgSample(
  occs = occs_Cp,
  bgMask =  bgMask_Cp,
  bgPtsNum = 100)
# Extract values of environmental layers for each background point
bgEnvsVals_Cp <- as.data.frame(raster::extract(bgMask_Cp,  bgSample_Cp))
##Add extracted values to background points table
bgEnvsVals_Cp <- cbind(scientific_name = paste0("bg_", "Carcharhinus perezi"), bgSample_Cp,
                            occID = NA, year = NA, institution_code = NA, country = NA,
                            state_province = NA, locality = NA, elevation = NA,
                            record_type = NA, bgEnvsVals_Cp)
```


### Partition occurrence data

Partition occurrences and background points for model training and
validation using “checkerboard”, a spatial partition method with an
aggregation factor of 2.

```{r}
# R code to get partitioned data
groups_Cp <- part_partitionOccs(
  occs = occs_Cp ,
  bg =  bgSample_Cp, 
  method = "cb1",
  bgMask = bgMask_Cp,
  aggFact = 2) 
```

### Build and Evaluate Niche Model

Generating a species distribution model using the maxnet algorithm as
implemented in ENMeval V2.0 (with clamping = FALSE). For tuning using L,
H feature classes and regularization multipliers in the 1, 2 range
increasing by 1. Not using any categorical predictor variables.

```{r}
# Run maxent model for the selected species
model_Cp <- model_maxent(
  occs = occs_Cp,
  bg = bgEnvsVals_Cp,
  user.grp = groups_Cp, 
  bgMsk = bgMask_Cp,
  rms = c(1, 2), 
  rmsStep =  1,
  fcs = c('L', 'H'),
  clampSel = FALSE,
  algMaxent = "maxnet",
  parallel = FALSE,
  numCores = 11)
```

### Visualize

Generate a map of the maxnet generated model with with a “p10” threshold
rule of 0.594019483220571.

```{r}
# Select current model and obtain raster prediction
m_Cp <- model_Cp@models[["fc.H_rm.1"]] 
predSel_Cp <- predictMaxnet(m_Cp, bgMask_Cp,
                                          type = "cloglog", 
                                          clamp = FALSE)
# extract the suitability values for all occurrences
occs_xy_Cp <- occs_Cp[c('longitude', 'latitude')]
# determine the threshold based on the current prediction
occPredVals_Cp <- raster::extract(predSel_Cp, occs_xy_Cp)
# Define probability of quantile based on selected threshold
thresProb_Cp <- switch("p10", 
                              "mtp" = 0, "p10" = 0.1, "qtp" = 0)
# Define threshold value
thres_Cp <- stats::quantile(occPredVals_Cp, probs = thresProb_Cp)
# Applied selected threshold
predSel_Cp <- predSel_Cp > thres_Cp

# Get values of prediction
mapPredVals_Cp <- getRasterVals(predSel_Cp, "cloglog")
# Define colors and legend  
rasCols <- c("#2c7bb6", "#abd9e9", "#ffffbf", "#fdae61", "#d7191c")
legendPal <- colorNumeric(rev(rasCols), mapPredVals_Cp, na.color = 'transparent')
rasPal <- c('gray', 'blue')
# Generate map
m <- leaflet() %>% addProviderTiles(providers$Esri.WorldTopoMap) 
m  %>%
  leaflet::addLegend("bottomright", colors = c('gray', 'blue'),
            title = "Thresholded Suitability<br>(Training)",
            labels = c("predicted absence", "predicted presence"),
            opacity = 1, layerId = "train") %>% 
  #add occurrence data
  addCircleMarkers(data = occs_Cp, lat = ~latitude, lng = ~longitude,
                   radius = 5, color = 'red', fill = TRUE, fillColor = "red",
                   fillOpacity = 0.2, weight = 2, popup = ~pop) %>% 
  ##Add model prediction
  addRasterImage(predSel_Cp, colors = rasPal, opacity = 0.7,
                 group = 'vis', layerId = 'mapPred', method = "ngb") %>%
 ##add background polygons
  addPolygons(data = bgExt_Cp, fill = FALSE,
              weight = 4, color = "blue", group = 'proj')
```

```{r}
# First must generate the transfer area according to the drawn polygon in the GUI
xfer_draw_Cp <-xfer_draw(
  polyXfXY = matrix(c(-81.169716, -81.125759, -83.675294, -85.653382, -84.994019, -83.015931, -81.169716, 21.655514, 23.320188, 23.763352, 23.320188, 21.205538, 21.164563, 21.655514),ncol=2,byrow=FALSE),
  polyXfID = 7698,
  drawXfBuf = 0)
# Create object of transfer variables
xferAreaEnvs_Cp <- envs_Cp
# Generate a transfer of the model to the desired area
xfer_area_Cp <- xfer_area(
  evalOut = model_Cp,
  curModel = "fc.H_rm.1",
  envs = xferAreaEnvs_Cp , 
  outputType = "cloglog",
  alg = "maxnet",
  clamp = FALSE,
  xfExt = xfer_draw_Cp) 
# store the cropped transfer variables
xferExt_Cp <- xfer_area_Cp$xferExt
 #map result
 
 ###Make map of transfer
bb_Cp <-  bgExt_Cp@bbox
bbZoom <- polyZoom(bb_Cp[1, 1], bb_Cp[2, 1], bb_Cp[1, 2], 
                   bb_Cp[2, 2], fraction = 0.05)
mapXferVals_Cp <- getRasterVals(xfer_area_Cp$xferArea,"cloglog")
rasCols_Cp <- c("#2c7bb6", "#abd9e9", "#ffffbf", "#fdae61", "#d7191c")
# if no threshold specified
legendPal <- colorNumeric(rev(rasCols_Cp), mapXferVals_Cp, na.color = 'transparent')
rasPal_Cp <- colorNumeric(rasCols_Cp, mapXferVals_Cp, na.color = 'transparent')
m <- leaflet() %>% addProviderTiles(providers$Esri.WorldTopoMap) 
m %>%
  fitBounds(bbZoom[1], bbZoom[2], bbZoom[3], bbZoom[4]) %>%
  leaflet::addLegend("bottomright", pal = legendPal,
            title = "Predicted Suitability<br>(Transferred)",
            values = mapXferVals_Cp, layerId = 'xfer',
            labFormat = reverseLabel(2, reverse_order = TRUE)) %>%
# map model prediction raster and transfer polygon
  clearMarkers() %>% clearShapes() %>% removeImage('xferRas') %>%
  addRasterImage(xfer_area_Cp$xferArea, colors = rasPal_Cp, opacity = 0.7,
                 layerId = 'xferRas', group = 'xfer', method = "ngb") %>%
 ##add transfer polygon (user drawn area)
  addPolygons(data = xfer_draw_Cp, fill = FALSE,
              weight = 4, color = "blue", group = 'xfer')
```

------------------------------------------------------------------------

## Analysis for *Stenella attenuata* (Sa)

### Obtain Occurrence Data

You searched the gbif database for *Stenella attenuata*, limited to 5000
records. You decided to remove occurrences without uncertainty
information? FALSE

```{r}
# Query selected database for occurrence records
queryDb_Sa <- occs_queryDb(
  spNames = "Stenella attenuata", 
  occDb = "gbif", 
  occNum = 5000,
  RmUncertain = FALSE)
occs_Sa <- queryDb_Sa$Stenella_attenuata$cleaned
```

### Obtain environmental data

Using user-specified variables.

```{r}
## Specify the directory with the environmental variables
dir_envs_Sa <- "D:/Wallace/GIT/MARSPECTca"
envs_path <- file.path(dir_envs_Sa, c('b1_EastWestAspect.tif', 'b2_NorthSouthAspect.tif', 'b3_PlanCurvature.tif', 'b4_ProfileCurvature.tif', 'b5_DistancetoShore.tif', 'b6_BathymetricSlope.tif', 'b7_Concavity.tif', 'b8_MeanAnnualSSS.tif', 'b9_MinimumMonthlySSS.tif', 'b10_MaximumMonthlySSS.tif', 'b11_AnnualRangeSSS.tif', 'b12_AnnualVarianceSSS.tif', 'b13_MeanAnnualSST.tif', 'b14_SSTcoldestIceFreeMonth.tif', 'b15_SSTwarmestIceFreeMonth.tif', 'b16_AnnualRangeSST.tif', 'b17_AnnualVarianceSST.tif', 'bathymetry_30s.tif'))
# Create environmental object 
envs_Sa <- envs_userEnvs(
  rasPath = envs_path,
  rasName = c('b1_EastWestAspect.tif', 'b2_NorthSouthAspect.tif', 'b3_PlanCurvature.tif', 'b4_ProfileCurvature.tif', 'b5_DistancetoShore.tif', 'b6_BathymetricSlope.tif', 'b7_Concavity.tif', 'b8_MeanAnnualSSS.tif', 'b9_MinimumMonthlySSS.tif', 'b10_MaximumMonthlySSS.tif', 'b11_AnnualRangeSSS.tif', 'b12_AnnualVarianceSSS.tif', 'b13_MeanAnnualSST.tif', 'b14_SSTcoldestIceFreeMonth.tif', 'b15_SSTwarmestIceFreeMonth.tif', 'b16_AnnualRangeSST.tif', 'b17_AnnualVarianceSST.tif', 'bathymetry_30s.tif'),
  doBrick = FALSE)
occs_xy_Sa <- occs_Sa[c('longitude', 'latitude')]
occs_vals_Sa <- as.data.frame(raster::extract(envs_Sa, occs_xy_Sa, cellnumbers = TRUE))
# Remove duplicated same cell values
occs_Sa <- occs_Sa[!duplicated(occs_vals_Sa[, 1]), ]
occs_vals_Sa <- occs_vals_Sa[!duplicated(occs_vals_Sa[, 1]), -1]
# remove occurrence records with NA environmental values
occs_Sa <- occs_Sa[!(rowSums(is.na(occs_vals_Sa)) >= 1), ]
# also remove variable value rows with NA environmental values
occs_vals_Sa <- na.omit(occs_vals_Sa)
# add columns for env variable values for each occurrence record
occs_Sa <- cbind(occs_Sa, occs_vals_Sa)
```

### Process Occurrence Data

Remove occurrences outside of user drawn polygon

```{r}
occs_Sa <- poccs_selectOccs(
  occs = occs_Sa,
  polySelXY = matrix(c(-78.576991, -82.31338, -82.13755, -78.313246, -77.653884, -76.071414, -75.016433, -73.653751, -71.236088, -67.455742, -63.103949, -61.521478, -61.433563, -62.708331, -63.147906, -63.499566, -64.290801, -78.576991, 17.012798, 14.942795, 10.442572, 10.010101, 9.143452, 10.053374, 11.521069, 11.133266, 13.32348, 11.693256, 11.607176, 12.681205, 14.177189, 16.25489, 17.096823, 18.018569, 17.976774, 17.012798), ncol = 2, byrow = FALSE),
  polySelID = 4817)
```

### Process environmental data

Sampling of 100 background points and corresponding environmental data
using a user drawn background extent with a 0 degree buffer.

```{r}
# Create a background extent based on user drawn polygon
bgExt_Sa <- penvs_drawBgExtent(
  polyExtXY = matrix(c(-80.071559, -82.665052, -82.665052, -81.170497, -78.401174, -78.168427, -77.662916, -76.135392, -75.739775, -75.080412, -74.497975, -73.662782, -72.585823, -71.607769, -69.434351, -68.060679, -65.115526, -62.56599, -61.51101, -61.693613, -62.572763, -64.111276, -68.023494, -71.408222, -74.79295, -78.177679, -79.452446, -80.071559, 16.928734, 14.815379, 11.262592, 9.75034, 9.793648, 9.426879, 9.144978, 9.719378, 10.573716, 11.134782, 11.231781, 11.382604, 12.038817, 12.671994, 12.822012, 11.705526, 10.951473, 11.350292, 13.656162, 16.043834, 17.51638, 17.976774, 17.264761, 17.432545, 17.264761, 17.264761, 17.432545, 16.928734),ncol=2,byrow=FALSE), 
  polyExtID = 6923, 
  drawBgBuf = 0, 
  occs = occs_Sa)
# Mask environmental data to provided extent
bgMask_Sa <- penvs_bgMask(
  occs = occs_Sa,
  envs = envs_Sa,
  bgExt = bgExt_Sa)
# Sample background points from the provided area
bgSample_Sa <- penvs_bgSample(
  occs = occs_Sa,
  bgMask =  bgMask_Sa,
  bgPtsNum = 100)
# Extract values of environmental layers for each background point
bgEnvsVals_Sa <- as.data.frame(raster::extract(bgMask_Sa,  bgSample_Sa))
##Add extracted values to background points table
bgEnvsVals_Sa <- cbind(scientific_name = paste0("bg_", "Stenella attenuata"), bgSample_Sa,
                            occID = NA, year = NA, institution_code = NA, country = NA,
                            state_province = NA, locality = NA, elevation = NA,
                            record_type = NA, bgEnvsVals_Sa)
```

### Partition occurrence data

Partition occurrences and background points for model training and
validation using “checkerboard”, a spatial partition method with an
aggregation factor of 2.

```{r}
# R code to get partitioned data
groups_Sa <- part_partitionOccs(
  occs = occs_Sa ,
  bg =  bgSample_Sa, 
  method = "cb1",
  bgMask = bgMask_Sa,
  aggFact = 2) 
```

### Build and Evaluate Niche Model

Generating a species distribution model using the maxnet algorithm as
implemented in ENMeval V2.0 (with clamping = FALSE). For tuning using L,
H feature classes and regularization multipliers in the 1, 2 range
increasing by 1. Not using any categorical predictor variables.

```{r}
# Run maxent model for the selected species
model_Sa <- model_maxent(
  occs = occs_Sa,
  bg = bgEnvsVals_Sa,
  user.grp = groups_Sa, 
  bgMsk = bgMask_Sa,
  rms = c(1, 2), 
  rmsStep =  1,
  fcs = c('L', 'H'),
  clampSel = FALSE,
  algMaxent = "maxnet",
  parallel = FALSE,
  numCores = 11)
```

### Visualize

Generate a map of the maxnet generated model with with a “p10” threshold
rule of 0.54760724560838.

```{r}
# Select current model and obtain raster prediction
m_Sa <- model_Sa@models[["fc.H_rm.1"]] 
predSel_Sa <- predictMaxnet(m_Sa, bgMask_Sa,
                                          type = "cloglog", 
                                          clamp = FALSE)
# extract the suitability values for all occurrences
occs_xy_Sa <- occs_Sa[c('longitude', 'latitude')]
# determine the threshold based on the current prediction
occPredVals_Sa <- raster::extract(predSel_Sa, occs_xy_Sa)
# Define probability of quantile based on selected threshold
thresProb_Sa <- switch("p10", 
                              "mtp" = 0, "p10" = 0.1, "qtp" = 0)
# Define threshold value
thres_Sa <- stats::quantile(occPredVals_Sa, probs = thresProb_Sa)
# Applied selected threshold
predSel_Sa <- predSel_Sa > thres_Sa

# Get values of prediction
mapPredVals_Sa <- getRasterVals(predSel_Sa, "cloglog")
# Define colors and legend  
rasCols <- c("#2c7bb6", "#abd9e9", "#ffffbf", "#fdae61", "#d7191c")
legendPal <- colorNumeric(rev(rasCols), mapPredVals_Sa, na.color = 'transparent')
rasPal <- c('gray', 'blue')
# Generate map
m <- leaflet() %>% addProviderTiles(providers$Esri.WorldTopoMap) 
m  %>%
  leaflet::addLegend("bottomright", colors = c('gray', 'blue'),
            title = "Thresholded Suitability<br>(Training)",
            labels = c("predicted absence", "predicted presence"),
            opacity = 1, layerId = "train") %>% 
  #add occurrence data
  addCircleMarkers(data = occs_Sa, lat = ~latitude, lng = ~longitude,
                   radius = 5, color = 'red', fill = TRUE, fillColor = "red",
                   fillOpacity = 0.2, weight = 2, popup = ~pop) %>% 
  ##Add model prediction
  addRasterImage(predSel_Sa, colors = rasPal, opacity = 0.7,
                 group = 'vis', layerId = 'mapPred', method = "ngb") %>%
 ##add background polygons
  addPolygons(data = bgExt_Sa, fill = FALSE,
              weight = 4, color = "blue", group = 'proj')
```

------------------------------------------------------------------------

## ESPACE analysis for *Carcharhinus perezi* and *Stenella attenuata* (Cp_Sa)

### Environmental space

Performing and plotting principal component analysis to reduce
dimensionality of environmental space for *Carcharhinus perezi* &
*Stenella attenuata*. PCA done for occsBg.

```{r}
# Determine the variables to use
pcaSel_Cp_Sa <- c('b1_EastWestAspect', 'b2_NorthSouthAspect', 'b3_PlanCurvature', 'b4_ProfileCurvature', 'b5_DistancetoShore', 'b6_BathymetricSlope', 'b7_Concavity', 'b8_MeanAnnualSSS', 'b9_MinimumMonthlySSS', 'b10_MaximumMonthlySSS', 'b11_AnnualRangeSSS', 'b12_AnnualVarianceSSS', 'b13_MeanAnnualSST', 'b14_SSTcoldestIceFreeMonth', 'b15_SSTwarmestIceFreeMonth', 'b16_AnnualRangeSST', 'b17_AnnualVarianceSST', 'bathymetry_30s')
# Run the pca
espace_pca_Cp_Sa <- espace_pca(
  sp.name1 = "Carcharhinus perezi",
  sp.name2 = "Stenella attenuata", 
  occs.z1 = occs_Cp[,pcaSel_Cp_Sa],
  occs.z2 = occs_Sa[,pcaSel_Cp_Sa],
  bgPts.z1 = bgEnvsVals_Cp[,pcaSel_Cp_Sa],
  bgPts.z2 = bgEnvsVals_Sa[,pcaSel_Cp_Sa])

## Generate plots
# PCA Scatter Plot
if ("occsBg" == "occs") {
  x <- espace_pca_Cp_Sa$scores[espace_pca_Cp_Sa$scores$bg == 'sp', ]
  x.f <- factor(x$sp)
} else if ("occsBg" == "occsBg") {
  x <- espace_pca_Cp_Sa$scores[espace_pca_Cp_Sa$scores$sp == 'bg', ]
  x.f <- factor(x$bg)
}
ade4::s.class(x, x.f, xax = 1, yax = 2,
              col = c("red", "blue"), cstar = 0, cpoint = 0.1)
title(xlab = paste0("PC", 1), ylab = paste0("PC", 2))
# PCA Correlation circle
ade4::s.corcircle(espace_pca_Cp_Sa$co, xax = 1, yax = 2,
                  lab = pcaSel_Cp_Sa, full = FALSE, box = TRUE)
      title(xlab = paste0("PC", 1),
            ylab = paste0("PC", 2))
# PCA screeplot
screeplot(espace_pca_Cp_Sa, main = NULL)
# Print PCA summary of results
summary(espace_pca_Cp_Sa)
```
